Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-12-29 WSSSPE2_report/WSSSPE2_review.tex
slight text change in 4.1, changed tables in 5 into figures both as requested by JORS reviewers

\documentclass[11pt, oneside]{amsart}

@@

Thaney spoke about Mozilla's work around code citation, through a collaboration
and prototype crafted between Mozilla, GitHub, figshare, and Zenodo. This work
was presented at a closed meeting in May 2014 at the National Institutes of
Health (NIH) around these issues, sparking a conversation from that meeting
around what a \emph{Software Discovery
Index}\footnote{Software Discovery Index, \url{http://softwarediscoveryindex.org/report/}} might look
like. The meeting included a number of publishers, researchers, and those behind
major scientific software efforts such as
Bioconductor\footnote{Bioconductor, \url{http://www.bioconductor.org}},
Galaxy\footnote{Galaxy, \url{http://galaxyproject.org}}, and
nanoHUB\footnote{nanoHUB, \url{https://nanohub.org}}.
%facilitates more efficient scientific research. SDI identifies scientific
%software by archiving and standardizing metadata for software and hence help
%connect both developers and users.
Ted Habermann in the audience commented that if the metadata is minimal, it
would be less onerous for data providers, but more burdensome for users---it
could be challenging to keep a balance between what have to be captured and what
would be ideal if we do not want to lose user engagement as in the case of the
old Harvard Dataverse, finding often only the first four fields of three pages
of metadata were filled out.

@@

The enablers of and barriers to software sustainability, roughly ranked by
attendee votes are shown in
{\tablename}s~\ref{tb:software_sustainability_enablers} and
\ref{tb:software_sustainability_barriers}, respectively.\footnote{A few other
items were suggested as barriers, but were not voted on due to lack of time in
the session:
layering up dependencies;
using software past its sustainable life;
using software past its usable life;
inertia for accepted answers versus wrong or right answers;
monolithic or poor code; and
need to restructure code when hardware/software/libraries change.
}
%\katznote{any discussion of these tables?}

\begin{table*}[ht]
\centering
\caption{Enablers of software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}
\label{tb:software_sustainability_enablers}
  \begin{small}
  \begin{tabular}{ | p{1.65cm} | p{12.0cm} |}
   %  \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
    \hline
  Importance & Item \\ \hline \hline
********** & healthy and vibrant communities; vibrant community to champion software \\ \hline
********** & designing for growth and extension -- open development \\ \hline
******* & culture in community for reuse \\ \hline
**** & portability \\ \hline
**** & culture in development community to support transition between developers \\ \hline
*** & interdisciplinary people: science + IT experience \\ \hline
** & planning for end of life \\ \hline
** & make smart choices about dependencies \\ \hline
* & thinking of software as product lines -- long term vs.~short term view \\ \hline
 & not all communities need new software \\ \hline
 & converting use into resources \\ \hline
    \end{tabular}
    \end{small}
\end{table*}

\begin{table}[ht]
\caption{Barriers to software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}
\label{tb:software_sustainability_barriers}
 \centering
  \begin{small}
  \begin{tabular}{ | p{1.65cm} | p{12.0cm} |}
% \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
   \hline
  Importance & Item \\ \hline \hline
******* & lack of incentives, including promotion and tenure process; promotion
and tenure process in academic is incompatible with sustainability \\ \hline
*****  & absent or poor documentation \\ \hline
***** & funding to ensure sustainability is difficult to obtain \\ \hline
**** & developers are not computer scientists; don't have software engineering
practices (in particular, those needed to scale-up projects to support and be
developed by a large sustainable community) \\ \hline
*** & overreliance on one or two people -- the `bus test'\footnote{``Bus test''
in Table~\ref{tb:software_sustainability_barriers} refers to the smallest number
of key people a project would need to lose to become non-viable (the larger the
number, the healthier the project).} \\ \hline
** & rate of change of underlying technologies \\ \hline
** & lack of business models for sustainability \\ \hline
* & lack of training for how to build sustainability into the system \\ \hline
 & maintenance for software is not visible, appears to ``just happen'' \\ \hline
 & licensing issues \\ \hline
 & staff turnover -- lack of continuity \\ \hline
\end{tabular}
 \end{small}
\end{table}

>>>>

\documentclass[11pt, oneside]{amsart}
\pdfoutput=1

@@

Thaney spoke about Mozilla's work around code citation, through a collaboration
and prototype crafted between Mozilla, GitHub, figshare, and Zenodo. This work
was presented at a closed meeting in May 2014 at the National Institutes of
Health (NIH) around these issues, sparking a conversation from that meeting
around what a \emph{Software Discovery
Index}\footnote{Software Discovery Index, \url{http://softwarediscoveryindex.org/report/}} might look
like. The meeting included a number of publishers, researchers, and those behind
major scientific software efforts such as
Bioconductor\footnote{Bioconductor, \url{http://www.bioconductor.org}},
Galaxy\footnote{Galaxy, \url{http://galaxyproject.org}}, and
nanoHUB\footnote{nanoHUB, \url{https://nanohub.org}}.
%facilitates more efficient scientific research. SDI identifies scientific
%software by archiving and standardizing metadata for software and hence help
%connect both developers and users.
Ted Habermann in the audience commented that if the metadata is minimal, it
would be less onerous for data providers, but more burdensome for users---it
could be challenging to keep a balance between what have to be captured and what
would be ideal if we do not want to lose user engagement, which he said he had
experienced with the
Harvard Dataverse~\cite{dataverse} where a large amount (three pages) of metadata
was requested, but often only the first four fields were filled out.

@@

The enablers of and barriers to software sustainability, roughly ranked by
attendee votes are shown in
%{\tablename}s~\ref{tb:software_sustainability_enablers} 
{\figurename}s~\ref{fig:software_sustainability_enablers}
and
%\ref{tb:software_sustainability_barriers}
\ref{fig:software_sustainability_barriers}, respectively.\footnote{A few other
items were suggested as barriers, but were not voted on due to lack of time in
the session:
layering up dependencies;
using software past its sustainable life;
using software past its usable life;
inertia for accepted answers versus wrong or right answers;
monolithic or poor code; and
need to restructure code when hardware/software/libraries change.
}
%\katznote{any discussion of these tables?}

%\begin{table*}[ht]
%\centering
%\caption{Enablers of software sustainability, with 0 to 10 `*'s roughly
%indicating the fraction of attendees who voted for an item as important.}
%\label{tb:software_sustainability_enablers}
%  \begin{small}
%  \begin{tabular}{ | p{1.65cm} | p{12.0cm} |}
%   %  \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
%    \hline
%  Importance & Item \\ \hline \hline
%********** & healthy and vibrant communities; vibrant community to champion software \\ \hline
%********** & designing for growth and extension -- open development \\ \hline
%******* & culture in community for reuse \\ \hline
%**** & portability \\ \hline
%**** & culture in development community to support transition between developers \\ \hline
%*** & interdisciplinary people: science + IT experience \\ \hline
%** & planning for end of life \\ \hline
%** & make smart choices about dependencies \\ \hline
%* & thinking of software as product lines -- long term vs.~short term view \\ \hline
% & not all communities need new software \\ \hline
% & converting use into resources \\ \hline
%    \end{tabular}
%    \end{small}
%\end{table*}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=1.0\textwidth]{figures/enablers}
  \caption{Enablers of software sustainability, showing the fraction of attendees (rounded to the nearest 10\%) who voted for an item as important.\label{fig:software_sustainability_enablers}}
\end{figure}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=1.0\textwidth]{figures/barriers}
  \caption{Barriers to software sustainability, showing the fraction of attendees (rounded to the nearest 10\%) who voted for an item as important.
{\footnotesize \newline (1) This item also included the fact that that the promotion and tenure process in academic is incompatible with sustainability.
\newline  (2) The software engineering practices were those in particular that are needed to scale-up projects to support and be developed by a large sustainable community).
\newline  (3) The smallest number of key people a project would need to lose to become
non-viable -- the larger the number, the healthier the project.}
\label{fig:software_sustainability_barriers}}
\end{figure}

%\begin{table}[ht]
%\caption{Barriers to software sustainability, with 0 to 10 `*'s roughly
%indicating the fraction of attendees who voted for an item as important.}
%\label{tb:software_sustainability_barriers}
% \centering
%  \begin{small}
%  \begin{tabular}{ | p{1.65cm} | p{12.0cm} |}
%% \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
%   \hline
%  Importance & Item \\ \hline \hline
%I & lack of incentives, including promotion and tenure process; promotion
%and tenure process in academic is incompatible with sustainability \\ \hline
%*****  & absent or poor documentation \\ \hline
%***** & funding to ensure sustainability is difficult to obtain \\ \hline
%**** & developers are not computer scientists; don't have software engineering
%practices (in particular, those needed to scale-up projects to support and be
%developed by a large sustainable community) \\ \hline
%*** & overreliance on one or two people -- the `bus test'\footnote{``Bus test''
%in Table~\ref{tb:software_sustainability_barriers} refers to the smallest number
%of key people a project would need to lose to become non-viable (the larger the
%number, the healthier the project).} \\ \hline
%** & rate of change of underlying technologies \\ \hline
%** & lack of business models for sustainability \\ \hline
%* & lack of training for how to build sustainability into the system \\ \hline
% & maintenance for software is not visible, appears to ``just happen'' \\ \hline
% & licensing issues \\ \hline
% & staff turnover -- lack of continuity \\ \hline
%\end{tabular}
% \end{small}
%\end{table}
